---
title: "Income and Ideology"
author: "Your Name Here"
date: "April 13, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(primer.data)
library(tidyverse)
library(rstanarm)
library(tidybayes)


# Check out the raw cces tibble, especially the faminc (family income) and
# ideology variables.
```

```{r}
x <- cces %>% 
  filter(year == 2019) %>% 
  select(state, gender, age, race, faminc, ideology) %>% 
  drop_na() %>% 
  
  # This is an absurd hack. Do not do this, ever. This only works because, if
  # you coerce factor variables into integers, the order of the factor levels is
  # preserved. So, low numbers for faminc/ideology correspond to poorer/liberal
  # individuals.
  
  # We need the variables, at least faminc, to be numeric so that we can create
  # a statistical model using the tools we have learned.
  
  mutate(faminc = as.integer(faminc)) %>% 
  mutate(ideology = as.integer(ideology))

table(cces$faminc)
table(x$faminc)
```


*Describe the relationship between ideology and family income. How does it vary by gender? Does your conclusion change if we "adjust" for race and age?*


## Wisdom
```{r}

x %>%
  ggplot(aes(x = ideology, y = faminc)) +
  geom_jitter(alpha = 0.05) +
  geom_smooth(formula = y ~ x,
              method = "lm")

# lower numbers of ideology are more liberal
# higher numbers of ideology are more conservative 

```


## Justice
What's the form of the model (linear or logistic)?
$y_i=\beta_0 + \beta_1x+ \varepsilon_i$

This one is linear.



## Courage

$ideology_i=\beta_0 + \beta_1faminc_i + \varepsilon_i$

```{r}
fit_1 <- stan_glm(data = x,
                  formula = ideology ~ faminc,
                  refresh = 0,
                  seed = 111)

print(fit_1, digits = 3)
```

$beta_0$ = intercept, when income is zero, estimate that ideology is 3.471 (slightly left of center)
3.471 is the median of the posterior distribution, 95% sure the real value is between +/- .05

$beta_1$ = slope, for every one unit increase in income (income bin of $10k), we would expect ideology to decrease slightly (approximately .047 more liberal)
*** This is a CAUSAL language - if I change a person's family income, then their ideology will change. That's the manipulation. 

When comparing two families that are all-else-equal, we would expect the richer family to be more liberal than the poorer family by a margin of -0.047.
*** This is PREDICTIVE language 

## Describe the relationship between ideology and family income. How does it vary by gender?

```{r}
fit_2 <- stan_glm(data = x,
                  formula = ideology ~ faminc + gender,
                  refresh = 0,
                  seed = 111)

print(fit_2, digits = 3)
```
$y_i = \beta_0 +\beta_1x_1i + \beta_2x_2i +\varepsilon_i$

Parallel Slope 


```{r}
fit_3 <- stan_glm(data = x,
                  formula = ideology ~ faminc + gender + (faminc * gender),
                  refresh = 0,
                  seed = 111)

print(fit_3, digits = 3)
```
$y_i = \beta_0 +\beta_1x + \beta_2x +\beta_3x_+\varepsilon_i$
$ideology_i = \beta_0 +\beta_1 faminc + \beta_2 gender +\beta_3 faminc * gender +\varepsilon_i$


*Does your conclusion change if we "adjust" for race and age?*
```{r}
fit_4 <- stan_glm(data = x,
                  formula = ideology ~ age + race + gender + faminc + (gender * faminc),
                  refresh = 0,
                  seed = 111)

print(fit_4, digits = 3)
```

Our interaction term did not change

## Temperance

Does vary by gender, controlling for race and age


```{r}
age <- 50 # just pick an age, 50 is near average
race <- "White" # just pick a race, white is the largest sample group
gender <- c("Male", "Female") # have to have both since that's the question - does it vary by income
faminc <- c(1, 6, 12) # simplify - no one right answer, just low, medium, and high income here

newobs <- expand_grid(age, race, gender, faminc)
# expand_grid creates all the combos of these things for us



pe <- posterior_epred(fit_4,
                      newdata = newobs) %>% 
  as_tibble() 
```

```{r}
z <- add_fitted_draws(newobs, fit_4)
```

```{r}
z %>%
  ggplot(aes(x = .value, 
             y = as.character(faminc), 
             fill = gender)) +
    
    stat_slab(alpha = 0.5) +
    labs(title = "Posterior for Ideology Varies by Income Bracket",
         subtitle = "Age at election does not matter to longevity for female candidates . . . ?", 
         x = "Ideology",
         y = "Income Bracket Level")
```

